---
title: "ERDR1 Colocalization"
output: html_document
date: "`r Sys.Date()`"
---

```{r htmlsetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r functions, include=FALSE}
library(tidyverse)
library(readxl)
library(MASS)
library(knitr)
library(kableExtra)
library(ggbeeswarm)
library(gridExtra)
library(ggpubr)
source(here::here("./Rscripts/colocalization_functions.R"))

```


```{r datasetup, include=FALSE, warning=FALSE, message=FALSE}

sex_analyze <- c("M", "F")
correct_p <- TRUE
plot_all_agg <- TRUE 
shownewplot <- TRUE 
showsections <- TRUE 
correctp <- TRUE 
toprocess <- "erdr1"

# key for calculations
erdr1_calc_key <- list(dapi_express_main = "main dapi / dapi",
                      tdt_express_spc = "tdt spc / tdt",
                      spc_express_tdt = "tdt spc / spc",
                      dapi_express_tdt_spc = "tdt spc / dapi",
                      dapi_express_tdt = "tdt / dapi",
                      dapi_express_spc = "spc / dapi",
                      tdt_express_main = "main dapi tdt / tdt",
                      tdt_express_main_spc = "main dapi tdt spc / tdt",
                      spc_express_main_tdt = "main dapi tdt spc / spc",
                      spc_express_main = "main dapi spc / spc"
                      )
                      


formlst <- list(erdr1=erdr1_calc_key)

# Note: to run this summary script, will need to provide this table of sample IDs
# columns should include: 'Day' (day sample was harvested), 'ID' (the numeric sample ID), 'GT' (sample genotype),
# 'Sex' (sex of sample), and 'BleoDose' (amount of bleomycin sample received) or 'Virus' (GT of AAV sample received)
mouseinfo <- read_xlsx(paste0(here::here(), "/mouse_id_info.xlsx"), sheet="ERDR1")
outputs_raw <- list.files(paste0(here::here(), "/out/erdr1/"), pattern="*.csv", recursive = T, full.names = T)

# separate out the additional sections to look at separately
outputs_all <- outputs_raw[!grepl("out_sep", outputs_raw)] 
outputs_all <- outputs_raw[!grepl("testing", outputs_raw)] 
outputs_all <- outputs_all[grep(toprocess, outputs_all)]

# separate out the additional sections to look at separately
outputs <- outputs_all[!grepl("addtl_sections", outputs_all)]

outs_sec1 <- createOuts(datfls=outputs, formnms=formlst, dapithresh=1000)
sumOuts_sec1 <- outs_sec1$split_section_output
slices <- 1  #1:3

sumOuts_list <- list(sumOuts_sec1)

# Process each section and bind the results
erdr1_df1 <- map2_dfr(sumOuts_list, slices, ~ processSection(.x, .y, "erdr1", mouseinfo, sex_analyze))

# set order
basenms_select <- c("section", "id", "type", "slice", "Day", "GT", "Sex")  #, "BleoDose")
if (nrow(erdr1_df1) > 0) {
  erdr1_df1 <- erdr1_df1 %>%
    dplyr::select(any_of(basenms_select),everything())
}


erdr1_start_idx <- min(grep('express', names(erdr1_df1)))

# erdr1
# change the 'main' names to 'erdr1'
names(erdr1_df1) <- gsub("main", "erdr1", names(erdr1_df1))
erdr1_calc_key <- lapply(erdr1_calc_key, function(x) gsub("main", "erdr1", x))
names(erdr1_calc_key) <- gsub("main", "erdr1", names(erdr1_calc_key))

erdr1_df1 <- erdr1_df1[,-which(names(erdr1_df1) == "Virus")]
# remove 'dapi' from names if dapi_colocalize is false
if (all(erdr1_df1$dapi_colocalize == FALSE)) {
  # remove dapi from names
  erdr1_calc_key <- lapply(erdr1_calc_key, function(x) {
    parts <- strsplit(x, "/", fixed = TRUE)[[1]]
  
    left  <- trimws(gsub("\\bdapi\\b", "", parts[1]))   # remove "dapi" ONLY on the left
    right <- trimws(parts[2])                           # preserve right side exactly
  
    paste(left, right, sep = " / ")
  })
}

erdr1_plt_out <- processDF(df=erdr1_df1, calc_key=erdr1_calc_key, start_idx=erdr1_start_idx, plotallagg=plot_all_agg,
                          show_p_correct=correctp, labelsize=20, ptsize=1.5, rm_outliers=F, makemainfig=T,
                          col_var="BleoType")
erdr1_df <- erdr1_plt_out$df

# Note: plot_out_X is a nested list of N long, where N is the number of parameters (erdr1 tdt spc / tdt, etc.) analyzed.
# For each of these, there is another list of 2 long, containing the plots and the wilcox test summaries.
# The plot length is 3
plot_vars_erdr1 <- erdr1_plt_out$plot_vars
plot_out_erdr1 <- erdr1_plt_out$plot_lst
plot_lst_erdr1 <- lapply(plot_out_erdr1, `[[`, 1)  # taking the 'plots' element of the list
plot_tst_erdr1 <- lapply(plot_out_erdr1, `[[`, 2)  # taking the 'tests' element of the list
dayGTtests_erdr1.l <- lapply(plot_tst_erdr1, `[[`, 1)
# full test df
dayGTtests_erdr1 <- bind_rows(dayGTtests_erdr1.l)
# check data in test
dayGTtests_erdr1.plots.l <- lapply(plot_tst_erdr1, `[[`, 2)
names(dayGTtests_erdr1.plots.l) <- sapply(plot_tst_erdr1, function(x) unique(x$wilcoxdf$param))


all_dfs <- list(erdr1=erdr1_df)

funckeys <- list(erdr1=erdr1_calc_key)

# if tests don't exist, set to NULL
# if (nrow(dayGTtests_erdr1) != 0) {
#   erdr1tst <- dayGTtests_erdr1
# } else {
  erdr1tst <- NULL
# }

```

#### Plotting percentages calculated per each 9x4 section with DAPI count > 0; SPC > 10; tdt > 10

# erdr1
## Analysis of `r ifelse(sex_analyze == "M", yes="Male", no=ifelse(sex_analyze == "F", yes="Female", no="NA"))` mice treated with `r unique(erdr1_df$BleoDose)` U/kg bleomycin

```{r info_erdr1, echo=FALSE, message=FALSE, eval=showsections}
if (!is.null(erdr1_df) & 'section' %in% names(erdr1_df)) {
  print(paste("IDs:", paste(unique(erdr1_df$id), collapse=", ")))
  print(paste("Sex:", paste(unique(erdr1_df$Sex), collapse=", ")))
  sectiondf1 <- erdr1_df %>% 
    group_by(id, slice) %>% 
    summarise(sections_used = paste(unique(section), collapse=", "))
  kable(sectiondf1)
}
```


#### Wilcoxon-Mann-Whitney test performed on 'slices combined' compare GT across mice per day
#### and GT per day between males and females, separately

```{r plot_main_erdr1AAA, fig.height=10, fig.width=19, echo=FALSE, warning=FALSE, message=FALSE, results='hide', eval=shownewplot}
if (!is.null(erdr1_df)) {
  # good dims: 6x17
  # new good dims: height=27, width=17
  # note: dimensions of plots are designed to view after knitting
  
  if (class(plot_lst_erdr1) != "try-error") {
    for (pltgrp in plot_lst_erdr1) {
      heights <- c(1, 2)
      newgrp <- pltgrp[-3]
      print(newgrp)
      cat('\n\n\n\n')
    }
  } else {
    print(paste("No data for sex", sex_analyze))
  }
}
```




### Wilcoxon-Mann-Whitney test
##### Table shows test of GT per day for 'M+F slices combined'
##### Bonferroni corrected across all `r length(unique(erdr1tst$Day))` day observations at alpha = 0.05
##### separately for each `r length(unique(erdr1tst$param))` parameters tested for erdr1 for GT per day
##### alpha~new~ = 0.05 / `r length(unique(erdr1tst$Day))` = `r 0.05 / length(unique(erdr1tst$Day))`


```{r erdr1test, echo=FALSE}
  if (!is.null(erdr1tst)) {
    kable(erdr1tst) %>% 
      kable_styling %>% 
      row_spec(which(erdr1tst$p_signif == 1), bold=T, hline_after=T, background="lightgrey") %>% 
      remove_column(which(names(erdr1tst) == "p_signif"))
  } else if (exists("erdr1_aov")) {  # not needed if not running anova
    kable(erdr1_aov) %>% 
      kable_styling %>% 
      row_spec(which(erdr1_aov$p < 0.05), bold=T, hline_after=T, background="lightgrey")
    kable(erdr1_aov) %>% 
      kable_styling %>% 
      row_spec(which(erdr1_aov$p < 0.05), bold=T, hline_after=T, background="lightgrey")
  } else {
    print("No comparison results for erdr1")
  }

```
